---
title: "Virginia Election Project"
author: "Hannah Loder"
execute:
  echo: true
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(kableExtra)
library(here)
options(scipen = 999)
options(stringsAsFactors = FALSE)
library(readxl) 
library(scales)
library(lubridate)
library(ggthemes)
library(esquisse)
library(tidyverse)
library(tigris)
library(sf)
library(tidycensus)
library(htmltools)
library(janitor)
library(here)
library(mapview)
library(leafsync)
library(leaflet.extras2)
options(tigris_class = "sf")

joined_vacomparison <- readRDS(here("joined_vacomparison.rds"))

```

```{r}
# cleaning up percentages 
joined_vacomparison <- joined_vacomparison %>% 
  mutate(youngkin_overperform = round_half_up(youngkin_overperform, 2),
         mcauliffe_underperform = round_half_up(mcauliffe_overperform, 2), 
         biden_overperform = round_half_up(biden_overperform, 2))
```

## Charting top 5 counties where Youngkin outperformed Trump:

Youngkin outperformed Trump in every Virginia county in the 2020 election. Below is the process I did for charting the top 5 counties where Youngkin performed better than Trump, meaning the places where the difference between their percentages of the vote was greatest.

```{r}
toptencounties_YO <- joined_vacomparison %>% 
  arrange(desc(youngkin_overperform)) %>% 
  head(5)
```

```{r}
ggplot(toptencounties_YO, aes(x = locality, y = youngkin_overperform)) +
  geom_col(color = "#B32B2B", fill = "#B32B2B") +
  coord_flip() +
  scale_y_continuous(name = "Percent Overperform", labels = scales::comma) +
  scale_x_discrete(name = "County") +
  labs(title = "Where Youngkin Outperformed Trump (%)", 
       subtitle = "Top 5 VA Counties 2020",
       caption = "VA Department of Elections") +
  theme_fivethirtyeight()
```

## Mapping Youngkin's Margin of Victory over Trump

Below is a map showing the difference in the percentage of votes Youngkin received in Virginia counties and those that Trump received. The map uses shading to show the degree to which Youngkin outperformed Trump in every county.

```{r}
 census_api_key("2a6f8c21a30d3024e038d67d7d4eba647dc79cd4", install=TRUE, overwrite=TRUE) 
```

```{r}
myvars <- c(totalpop = "B01003_001",
            medincome = "B19013_001",
            medage = "B01002_001"
)
```

```{r}
va_counties_withgeo <- get_acs(geography = "county",
                       variables = c(myvars),
                       state = "VA",
                       output = "wide",
                       geometry = TRUE)

va_counties_withgeo
```

```{r}
va_counties_withgeo <- va_counties_withgeo %>% 
  mutate(
    NAME = str_replace(NAME, ", Virginia", ""),
    NAME = str_to_upper(NAME)
  )
```

```{r}
joined <- left_join(va_counties_withgeo, joined_vacomparison, by = c("NAME" = "locality"))
```

```{r}

mylabel <- glue::glue("{joined$NAME} {joined$youngkin_overperform}")

mypopup <- glue::glue("<strong>{joined$NAME}</strong><br />
                      Percent: {joined$youngkin_overperform}<br />
                      Total Population: {joined$totalpopE}" ) %>% 
  lapply(htmltools::HTML)

mapview(joined, zcol = "youngkin_overperform", 
         col.regions = RColorBrewer::brewer.pal(9, "Reds"), 
         alpha.regions = 1,
         popup = mypopup, label = mylabel)

```

## Charting top 5 counties where Biden outperformed McAuliffe:

McAuliffe did not outperform Biden in any county, as referenced in the data. For this reason, I decided to chart the top ten counties where Biden outperformed McAuliffe.

```{r}
toptencounties_BO <- joined_vacomparison %>% 
  arrange(desc(biden_overperform)) %>% 
  head(5)
```

```{r}
ggplot(toptencounties_BO, aes(x = locality, y = biden_overperform)) +
  geom_col(color = "#287AC7", fill = "#287AC7") +
  coord_flip() +
  scale_y_continuous(name = "Percent Overperform", labels = scales::comma) +
  scale_x_discrete(name = "County") +
  labs(title = "Where Biden Outperformed McAuliffe (%)", 
       subtitle = "Top 5 VA Counties 2020",
       caption = "VA Department of Elections") +
  theme_fivethirtyeight()
```

## Mapping Biden's Margin of Victory over McAuliffe

I mapped the difference in the percentage of votes Biden received compared to those McAuliffe received. Biden outperformed McAuliffe in every Virginia county. The darker the shading, the greater the victory.

```{r}
map2label <- glue::glue("{joined$NAME} {joined$biden_overperform}")

popupmap2 <- glue::glue("<strong>{joined$NAME}</strong><br />
                      Percent: {joined$biden_overperform}<br />
                      Total Population: {joined$totalpopE}" ) %>% 
  lapply(htmltools::HTML)

mapview(joined, zcol = "biden_overperform", 
         col.regions = RColorBrewer::brewer.pal(9, "Blues"), 
         alpha.regions = 1,
         label = map2label, popup = popupmap2)
```

## Charting Where Youngkin Got the Highest Percentage of the Vote

Youngkin won the Virginia Governor race in 2020. Below is the code I used to create a chart showing the five counties in Virginia where Youngkin got the highest percentage of the vote.

```{r}
youngkin_top5 <- joined_vacomparison %>% 
  arrange(desc(pct_youngkin)) %>% 
  head(5)
```

```{r}
ggplot(youngkin_top5, aes(x = locality, y = pct_youngkin)) +
  geom_col(color = "#FF3636", fill = "#FF3636") +
  coord_flip() +
  scale_y_continuous(name = "Percent Youngkin", labels = scales::comma) +
  scale_x_discrete(name = "County") +
  labs(title = "Youngkin Percent of Votes", 
       subtitle = "Top 5 VA Counties 2020",
       caption = "VA Department of Elections") +
  theme_fivethirtyeight()
```

## Charting Where McAuliffe Got the Highest Percentage of the Vote

While McAuliffe did not won the Virginia Governor race, he did receive the highest percentage of the vote in some counties. I created a chart showing the top five counties in Virginia where McAuliffe got the highest percentage of the vote.

```{r}
mcauliffe_top5 <- joined_vacomparison %>% 
  arrange(desc(pct_mcauliffe)) %>% 
  head(5)
```

```{r}
ggplot(mcauliffe_top5, aes(x = locality, y = pct_mcauliffe)) +
  geom_col(color = "#4CA0FF", fill = "#4CA0FF") +
  coord_flip() +
  scale_y_continuous(name = "Percent McAuliffe", labels = scales::comma) +
  scale_x_discrete(name = "County") +
  labs(title = "McAuliffe Percent of Votes", 
       subtitle = "Top 5 VA Counties 2020",
       caption = "VA Department of Elections") +
  theme_fivethirtyeight()
```

## Mapping Voting Percentages of Youngkin and McAuliffe

Below is a map showing the percentages of the vote Youngkin and McAuliffe received in Virginia counties. The maps of each of their performances are joined together to make one interactive map.

```{r}

youngkin_label <- glue::glue("{joined$NAME} {joined$pct_youngkin}")

popup_youngkin <- glue::glue("<strong>{joined$NAME}</strong><br />
                      Percent Youngkin: {joined$pct_youngkin}<br />
                      Total Population: {joined$totalpopE}" ) %>% 
  lapply(htmltools::HTML)


mcauliffe_label <- glue::glue("{joined$NAME} {joined$pct_mcauliffe}")

popup_mcauliffe <- glue::glue("<strong>{joined$NAME}</strong><br />
                      Percent McAuliffe: {joined$pct_mcauliffe}<br />
                      Total Population: {joined$totalpopE}" ) %>% 
  lapply(htmltools::HTML)

map_youngkin <- mapview(joined, zcol = "pct_youngkin", 
         col.regions = RColorBrewer::brewer.pal(9, "Reds"), 
         alpha.regions = 1, 
         label = youngkin_label, popup = popup_youngkin)

map_mcauliffe <- mapview(joined, zcol = "pct_mcauliffe", 
         col.regions = RColorBrewer::brewer.pal(9, "Blues"), 
         alpha.regions = 1, 
         label = mcauliffe_label, popup = popup_mcauliffe)

map_mcauliffe | map_youngkin

```
